A fault tolerant multi-server system for MMOGs

### 2010-11-10 textos/phd-prospectus/prospectus.tex

all the mmog intro and context...

@@

The next sections will describe the services the service must provide to the players connected to it (Section \ref{sec:services}), the specific requirements for a fault-tolerant distributed MMOG server system, considering the crash-stop failure model, and the proposed distributed architecture (Section \ref{sec:feats}) and an algorithm devised to synchronize the game state between the replicas (Section \ref{sec:algorithm}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Characteristics and architecture of the intended distributed game server system} \label{sec:feats}

@@

However, several questions arise when dealing with a potentially large scale geographically distributed server system composed of voluntary nodes. First, as the nodes are volunteer, there is no guarantee that they will be available whenever they are needed, neither there is any guarantee regarding the amount of resources that each of these nodes makes available for the system -- be it cpu power, network bandwidth or storage. Besides, not only can these nodes become unavailable, but there can be a high rate of nodes joining and leaving the system, characterizing churn. There are also problems related to the geographical distribution of this system, as the option of using efficient group communication techniques -- such as network level multicast -- is not widely available on the Internet. Finally, considering the potentially large number of nodes in this system, it is very likely that some of them present failures, for what we are assuming the crash-stop failure model -- therefore, we assume the possibility of nodes crashing and network links becoming unavailable.

%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Proposed architecture} \label{sec:arch}

%x

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proposed synchronization algorithm} \label{sec:algorithm}

y

>>>>

all the mmog intro and context goes here...

@@

The next sections will describe the services the service must provide to the players connected to it (Section \ref{sec:services}), the specific requirements for a fault-tolerant distributed MMOG server system, considering the crash-stop failure model (Section \ref{sec:feats}), and the proposed distributed architecture (Section \ref{sec:arch}) and an algorithm devised to synchronize the game state between the replicas (Section \ref{sec:algorithm}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Characteristics of the intended distributed game server system} \label{sec:feats}

@@

However, several questions arise when dealing with a potentially large scale geographically distributed server system composed of voluntary nodes. First, as the nodes are volunteer, there is no guarantee that they will be available whenever they are needed, neither there is any guarantee regarding the amount of resources that each of these nodes makes available for the system -- be it cpu power, network bandwidth or storage. Besides, not only can these nodes become unavailable, but there can be a high rate of nodes joining and leaving the system, characterizing churn. There are also problems related to the geographical distribution of this system, as the option of using efficient group communication techniques -- such as network level multicast -- is not widely available on the Internet.

Finally, considering the potentially large number of nodes in this system, it is very likely that some of them present failures, for what we are assuming the crash-stop failure model -- therefore, we assume the possibility of nodes crashing and network links becoming unavailable. The services provided by the system must continue even in the presence of such failures. Not only must the distribution itself be transparent to the players, but also the existence of failures and their countermeasures must not be noticeable by them.

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proposed architecture} \label{sec:arch}

In order to split a game server into several smaller units executed by many volunteer nodes, it is used the idea of dividing the virtual environment in regions, each of which may be managed by a different server -- or set of server. There are many approaches proposed in the literature, such as using fixed size cells which are grouped into regions, or using binary space partition trees, and so on. For the sake of generality, the architecture proposed here  does not decide how exactly the partitioning of the virtual environment will be performed.

One of the assumptions here is that the events of the game lead to a deterministic resulting game state. In other words, if the game starts in the state $S_{a}$, there is only one possible state $S_{b}$ to be reached after a sequence $E=\{e_{1}, e_{2}, ..., e_{n}\}$ of events. So, the game state may be modeled as a state machine, which is replicated across the different game server nodes. These server nodes are divided in two categories: \textbf{region managers} and \textbf{region replicas}. Each region has only one manager, who is responsible for ordering the events that happen inside that region. The region replicas have no authority to order the events -- be it players' commands or non-player events, such as the actions of AI-controlled characters --, but they can deliver and process any event $e$, as long as the game state in their memory results from every game state \textit{causally} prior to $e$. We later present the causality definition considered here.

As proposed in [cite related work], it is used here the idea of \textbf{contact servers} and \textbf{target servers}. In our architecture, the players whose avatars are in a certain region are distributed among the server nodes responsible for that region. The distribution is decided based on the network latency for each pair client-server. Each client is assigned to a server to which it has the lowest network latency among all servers in its region. This server, to which the client communicates directly, is its contact server. The target server is always the region manager, which can also be the contact server for certain players. When one of the servers gets overloaded because of the broadcasting of messages to players, these messages are then sent via other routes, even if the network latency of that new route is higher. Figure \ref{fig:arch} illustrates this architecture.

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.45\linewidth]{images/arch}
	\caption{Basic architecture with Region Managers and Region Replicas}
	\label{fig:arch}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proposed causality definition and synchronization algorithm} \label{sec:algorithm}

When a player issues a command $c_{i}$, it is sent to its contact server, which forwards the command to the region manager. The region manager then processes the command sent by the player and calculates its resulting state, also considering every command -- or any kind of event -- $c_{j}$ sent by other players, such that $c_{j}$ was issued prior to $c_{i}$. Any prior event $c_{j}$ that the contact server should be aware of and that has not been received by the contact server of that player yet is then sent by the region manager. By having these prior events, the contact server is able to calculate the new game state that should be sent to the player who issued the command $c_{i}$.

The \textbf{causality} is defined here based on which objects are affected by an event. If commands $e_{i}$ and $e_{j}$ affect the same object, then they must be ordered. The order is defined by the time when they are delivered at the region manager responsible for that object.

asdf

We could also use the other idea (player --tob--> region servers; then region servers --lazy-synch-->all other servers)
